Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India : vertical stratification reflects guild-depedent dynamics

### 2016-08-08 inst/doc/draft.Rmd
v0.1.9029 Corrected intro

---
title: Functional assembly into and among community of a tropical evergreen forest
  from Western Ghats, India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: word_document
bibliography: refs.bib
---

@@

functional trait; tropical evergreen forest; Western Ghats; India...

@@

Ecology offers different theories describing community composition of tropical forest. Niche theory invoke ecological filtering responsible for species assembly inside communities [@chave_comparing_2002]. On the other hand neutral theory [@hubbell_neutral_2005] assume functional equivalence of species, implying species distributions shaped by stochastic events as dispersion and death. @lortie_rethinking_2004 offered a conceptual framewok that reconciliates the two approach by defining integrated community concept. Integrated communities are build on three levels: $\gamma$ diversity as the initial regional species pool filtered by stochastic neutral events, $\beta$ diversity between communities resulting from abiotic filtering and stochastic events among communities, and $\alpha$ diversity defining the community assembly after local abiotic and biotic filtering.

Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction. Functional traits can thus be linked to trade-offs representative of species ecological strategies. For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary in so called economic spectra. @wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009. The leaf economic spectrum opposes productive ecological strategy to conservative persistent species. The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. A debated question is links between species economic spectra and strategy axes. Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones. Stochastic process and environmental filtering will be highly scale dependent. Effectively some functional traits will be more variable at the individual level with plasticity to micro-environmental conditions whereas others will be more affected on global environmental gradient with phylogeny. For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.

Functional traits in community ecology are linking ecological strategies, community assembly theory and functional diversity [@westoby_land-plant_2006]. In this paper we present the use of traits-based approach combined to community ecology to encompass functional assembly into and among community of a tropical evergreen forest from Western Ghats, India. Few studies investigated functional ecology whithin the Western Ghats indian biodiversity hotspot [@myers_biodiversity_2000]. Western Ghats present atypic moist tropical evergreen forest though seasonnaly dry and afffected by strong monsoon events. Such environmental conditions can be highly interesting to discuss and evaluate community assembly theory resulting in functional assembly of those forests. To do so we organised our study framework in two main questions :

@@

The study was carried out in the the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest. The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot. Climate is characterized by a peak of heavy rainfall during monsoon and a marked dry season. Annual precipitation averages 5108 mm with 90% between June and October and an average dry season (monthly rainfall $\leq 45$ mm) of 3.8 months [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998]. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) were monitored in an area progressively extended to 10ha in 2014. All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years.

>>>>

---
title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India : vertical stratification reflects guild-depedent dynamics'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

@@

functional trait; rainforest; dry season; forest strata; Western Ghats; ...

@@

Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975]. Tropical rainforests have especially fascinated standing issue in ecology [@connell_diversity_1978]. Two concurrent theories have been proposed to explain how species coexistence can be maintained in ecological communities. Niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001]. Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations. 
These approaches have been applied to resolve the conundrum og high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce. We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where no study has addressed the relative influence of neutral and niche-based dynamics on local forest composition.

The variation of functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007]. Thes variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002]. For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004] and wood [@chave_towards_2009] economic spectra. The leaf economic spectrum opposes productive ecological strategy to conservative persistent strategy. The wood economic spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977: (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. @baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strtegies, but @reich_world-wide_2014 conversely advocated that the two specrta have to integrated in a comprehensive whole plant economic spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones. Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest. Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004]. Determining the extent of neutral and niche-based dynamics then implies (i) decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and (ii) designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012]. We will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models.

Apart from scale-dependent dynamics in space whithin and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009]. The neutral theory has been basically applied in rainforests to adress diversity patterns in the guild of canopy trees [@hubbell_unified_2001], thus ignoring dynamics of understorey trees. However basic ecological constraints drive resource use strategies and biotic interactions in this ecosystems and imply considering functional strategies and relationship among guilds as well. Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988]. It thus provide an ideal context to analyze guilod-dependent dynamics and ecological strategies in tropical rainforests. We will thus address diversity patterns and community assembly at guild-level.

In this paper, we used a trait-based approach to adress guild and community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India. Although of wet evergreen type, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon. In this context, we expect water stress and competition for resource to be key drivers of forest dynamics. We thus adressed two main questions:

@@

The study was carried out in the 10-ha "Uppangala plot" of undisturbed, old-growth wet evergreen Diptrocarp forest. The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India) in India's Western Ghats biodiversity hotspot. Climate is characterized by a peak of heavy rainfall during monsoon and a marked dry season. Annual precipitation averages 5108 mm with 90% between June and October and an average dry season (monthly rainfall $\leq 45$ mm) of 3.8 months [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topography of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the major local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998]. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. Since 1989, all trees $\geq 30$ cm girth at breast height (gbh) were monitored in an area progressively extended to 10ha in 2014. All trees were identified, tagged, mapped, and provided with permanent dendrometer bands censused every 2-5 years.
